### 2017-10-21 RevisedSubmission/hines_hetland_2017_gji.tex
justified text file

\documentclass[extra,mreferee]{gji}
%\documentclass[extra,onecolumn,doublespacing]{gji}
\usepackage{timet}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{arydshln}
\usepackage{bm}
\usepackage{ulem}

\title[Transient strain in geodetic data]
      {Revealing transient strain in geodetic data with Gaussian process regression}

\author[T. T. Hines and E. A. Hetland]
       {T. T. Hines$^1$ and E. A. Hetland$^1$ \\
        $^1$ Department of Earth and Environmental Sciences, University of Michigan, Ann Arbor, MI, USA}

\date{Received XXX; in original form XXX}

\pagerange{\pageref{firstpage}--\pageref{lastpage}}

\volume{XXX}

\pubyear{XXX}

% Authors with AMS fonts and mssymb.tex can comment out the following
% line to get the correct symbol for Geophysical Journal International.
\let\leqslant=\leq

\newtheorem{theorem}{Theorem}[section]

\begin{document}

\label{firstpage}

\maketitle

%% DEFINE MATH MODE MACROS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\pos}{\vec{x}} % position vector
\newcommand{\data}{\mitbf{d}} % data vector
\newcommand{\post}{\hat{u}} % posterior
\newcommand{\points}{\mitbf{P}} % observation points
\newcommand{\strain}{\dot\varepsilon} % strain rate
\newcommand{\E}[1]{\e\left[ #1 \right]} % Expected value
\newcommand{\Cov}[1]{\mathrm{Cov}\left[ #1 \right]} % Expected value
\newcommand{\e}{\mathrm{e}} % east subscript
\newcommand{\n}{\mathrm{n}} % north subscript
\newcommand{\G}{\mitbf{G}} % basis functions
\newcommand{\zeros}{\mitbf{0}} % zeros matrix
\newcommand{\eye}{\mitbf{I}} % identity matrix

%% SUMMARY
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{summary}

Transient strain rates derived from GNSS data can be used to detect
and understand geophysical phenomena such as slow slip events and
postseismic deformation. Here we propose using Gaussian process
regression (GPR) as a tool for estimating transient strain rates from
GNSS data. GPR is a non-parametric, Bayesian method for interpolating
scattered data. Transient strain rates estimated with GPR have
meaningful uncertainties, allowing geophysical signal to be easily
discerned from noise. In our approach, we assume a stochastic prior
model for transient displacements. The prior describes how much one
expects transient displacements to covary spatially and temporally. A
posterior estimate of transient strain rates is obtained by
differentiating the posterior displacements. One limitation with GPR
is that it is not robust against outliers, so we introduce a
pre-processing method for detecting and removing outliers in GNSS
data. As a demonstration, we use GPR to detect transient strain
resulting from slow slip events in the Pacific Northwest. Maximum
likelihood methods are used to constrain a prior model for transient
displacements in this region. The temporal covariance of our prior
model is described by a compact Wendland covariance function, which
significantly reduces the computational burden that can be associated
with GPR. Our results reveal the spatial and temporal evolution of
strain from slow slip events. We verify that the transient strain
estimated with GPR is in fact geophysical signal by comparing it to
the seismic tremor that is associated with Pacific Northwest slow slip
events.

\end{summary}

\begin{keywords}
 XXX -- XXX -- XXX -- XXX.
\end{keywords}

%% INTRODUCTION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}\label{sec:Introduction}

Crustal strain rates are fundamentally important quantities for
assessing seismic hazard. Knowing where and how quickly strain is
accumulating gives insight into where we can expect stored elastic
energy to be released seismically. Consequently, secular crustal
strain rates estimated from GNSS data have been used to constrain
seismic hazard models such as UCERF3 \citep{Field2014}. Transient
crustal strain, which is caused by geophysical phenomena such as slow
slip events (SSEs) or postseismic deformation, is also relevant for
assessing seismic hazard. While transient strain itself is not
damaging, there is a risk that it can trigger major earthquakes
\citep{Roeloffs2006,Freed2001}. Dense networks of continuous GNSS
stations, such as the Plate Boundary Observatory (PBO), make it
possible to identify transient strain with high fidelity. Developing
and improving upon methods for deriving secular and transient strain
from GNSS data is an active area of research.

Most methods for estimating strain rates from GNSS data assume some
parametric form for the deformation signal. The simplest method for
estimating secular strain rates assumes that GNSS derived velocities
can be described with a first-order polynomial (i.e, the deformation
gradients are constant) over some subnetwork of the GNSS stations
\citep[e.g.,][]{Feigl1990,Murray2000}. The components of the strain
rate tensor for each subnetwork are then determined from the least
squares fit to the observations. The assumption that deformation
gradients are spatially uniform is not appropriate when subnetworks
span too large of an area. To help overcome this deficiency,
\citet{Shen1996,Shen2015} used an inverse distance weighting scheme,
in which the estimated strain rate at a point is primarily controlled
by observations at nearby stations. The method of
\citet{Shen1996,Shen2015} can be viewed as a form of local least
squares regression with a first-order polynomial \citep[e.g.,][sec.
6]{Hastie2009}. Other methods for estimating secular strain rates have
parameterized GNSS derived velocities with bi-cubic splines
\citep{Beavan2001}, spherical wavelets \citep{Tape2009}, and
elastostatic Green's functions \citep{Sandwell2016}. The type of basis
functions and the number of degrees of freedom for a parameterization,
which are often chosen subjectively, have a strong influence on the
strain solution. If there are too few degrees of freedom in the
parameterization, then estimated strain rates will be biased and the
uncertainties will be underestimated. On the other hand, if there are
too many degrees of freedom, then there will not be any coherent
features in the estimated strain rates. The methods described by
\citet{Beavan2001} and \citet{Tape2009} also require the user to
specify penalty parameters that control a similar trade-off between
bias and variance in the solution. One could parameterize deformation
with a physically motivated model of interseismic deformation
\citep[e.g.,][]{Meade2005,McCaffrey2007}. In such models the
lithospheric rheology and fault geometries are assumed to be known.
Any errors in the assumed physical model could result in biased strain
estimates and underestimated formal uncertainties.

The aforementioned studies are concerned with estimating secular
strain rates. In recent years the Southern California Earthquake
Center (SCEC) community has shown interest in developing methods for
detecting transient strain. SCEC supported a transient detection
exercise \citep{Lohman2013}, where several research groups tested
their methods for detecting transient geophysical signal with a
synthetic GNSS dataset. Among the methods tested were the Network
Strain Filter (NSF) \citep{Ohtani2010} and the Network Inversion
Filter (NIF) \citep{Segall1997}. The NSF uses a wavelet
parameterization to describe the spatial component of geophysical
signal. The NIF, which is intended for imaging slow fault slip from
geodetic data, uses the elastic dislocation Green's functions from
\citet{Okada1992}. For the NSF and NIF, the time dependence of the
geophysical signal is modeled as integrated Brownian motion. The
method described in \citet{Holt2013} was also tested in the SCEC
transient detection exercise, which calculates strain rates using a
bi-cubic spatial parameterization of displacements between time
epochs. \citet{Holt2013} defined a detection threshold based on the
strain rate magnitude, and below we demonstrate that this is indeed an
effective criterion for identifying geophysical signal. For the same
reasons descibed above, the transient deformation and corresponding
uncertainties estimated by these methods can be biased by the chosen
spatial parameterization. It is then difficult to distinguish signal
from noise with these methods, which limits their utility for
transient detection.

Here we propose using Gaussian process regression (GPR)
\citep{Rasmussen2006} to estimate transient strain from GNSS data. GPR
is closely related to kriging \citep{Cressie1992} and least squares
collocation \citep{Moritz1978}. The latter has been used by
\citet{Kato1998} and \citet{El-Fiky1999} to estimate secular strain
rates from GNSS data. GPR is a Bayesian, non-parametric method for
inferring a continuous signal from scattered data. Since GNSS stations
are irregularly spaced and observation times may differ between
stations, GPR is an ideal tool for synthesizing discrete GNSS data
into a spatially and temporally continuous representation of surface
deformation. GPR is Bayesian in that we describe our prior
understanding of the geophysical signal with a Gaussian process. A
Gaussian process is a continuous stochastic process whose value at any
collection of points can be described as a Gaussian random vector.
Just as a Gaussian random vector is fully determined by a mean vector
and a covariance matrix, a Gaussian process is fully determined by a
mean function and a covariance function. For example, Brownian motion,
$B(t)$, is a well known Gaussian process in $\mathbb{R}^1$ which has
the mean function $\E{B(t)} = 0$ and the covariance function
$\Cov{B(t),B(t')} = \min(t,t')$, where $t,t' \ge 0$. If there is no
information available to help choose an appropriate prior Gaussian
process, then maximum likelihood methods can be used to objectively
choose one that is most consistent with the data. Once a prior is
selected, we incorporate GNSS observations with the prior to form a
posterior estimate of transient strain. The posterior transient strain
is also a Gaussian process, and we can use its distribution to
confidently discern geophysical signal from noise. We use GPR to infer
strain resulting from SSEs in the Pacific Northwest, demonstrating
that GPR is an effective tool for detecting transient geophysical
processes.

%% ESTIMATING TRANSIENT STRAIN RATES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Estimating Transient Strain Rates}\label{sec:Method}

We seek a spatially and temporally dependent estimate of transient
strain rates. We consider transient strain rates to be any deviation
from secular strain rates. We denote transient strain rates as
$\strain(p)$, where $p$ represents the ordered pair $(\pos,t)$, $\pos
= (x_\e,x_\n)$ are spatial coordinates, and $t$ is time. For
simplicity, we assume that the study region is sufficiently small that
$\pos$ can be considered a point in a 2-D map projection that is
aligned with the cardinal directions. We determine $\strain$ by
spatially and temporally differentiating estimates of transient
displacements, which are constrained by GNSS data.

We let $\vec{u}(p) = (u_\e(p),u_\n(p))$ be our prior understanding of
transient displacements. Since transient displacements are not
precisely known, we do not consider $\vec{u}$ to be a deterministic
function. Instead, $\vec{u}$ is a stochastic process containing a
distribution of functions that could potentially describe transient
displacements. We let each component of $\vec{u}$ be a Gaussian
process with zero mean and a generic covariance function
$\Cov{u_i(p),u_i(p')} = C_{u_i}(p,p')$. Using a more concise notation,
we write our prior on each component of $\vec{u}$ as $u_i \sim
\mathcal{GP}\left(0,C_{u_i}\right)$. The function $C_{u_i}$ must be
positive definite in order to be a valid covariance function. By
definition, $C_{u_i}$ is a positive definite function if the matrix
$[C_{u_i}(p,p')]_{(p,p') \in \points \times \points}$ is positive
definite for any set of points $\points$ \citep[sec.
2.5]{Cressie1992}. We assume that $C_{u_i}$ can be separated into
spatial and temporal functions as
\begin{equation}\label{eq:TransientCovariance}
C_{u_i}(p,p') = C_{u_i}\left((\pos,t),(\pos',t')\right) 
              = X_i(\pos,\pos')T_i(t,t').
\end{equation}   
As long as the functions $X_i$ and $T_i$ are positive definite,
$C_{u_i}$ is guaranteed to also be positive definite \citep[sec.
4.2.4]{Rasmussen2006}. We also require that the derivatives
$\partial^2 X_i(\pos,\pos')/ \partial x_j \partial x_j'$ and
$\partial^2 T_i(t,t') / \partial t \partial t'$ exist. This ensures
that $\vec{u}$ is spatially and temporally differentiable, allowing us
to compute $\strain$ (See \citet[sec. 2.2]{Adler1981} or \citet[sec.
10A]{Papoulis1991} for a definition of stochastic differentiation and
the conditions for differentiability).

To give an example of $u_i$, we can use a squared exponential function
for $X_i$ and $T_i$,
\begin{equation}\label{eq:ExampleCov}
X_i(\pos,\pos') = \exp\left(\frac{-||\pos - \pos'||_2^2}{2\ell^2}\right), \ \ \ \ 
T_i(t,t') = \phi^2 \exp\left(\frac{-|t - t'|^2}{2\tau^2}\right),
\end{equation}
which satisfies our requirements for positive definiteness and
differentiability. The parameters $\ell$ and $\tau$ control the
length-scale and time-scale, respectively, of realizations (i.e.,
sample functions) of $u_i$. The parameter $\phi$, which we have
arbitrarily chosen to incorporate into $T_i$ rather than $X_i$,
controls the amplitude of realizations of $u_i$. Ideally, we want
realizations of $u_i$ to have a length-scale, time-scale, and
amplitude that resembles what we expect from true signal. A few
realizations for $u_i$ are shown in Figure XXX. While the squared
exponential function is a commonly used covariance function for GPR,
it is not appropriate for every application. The appropriate choice
for $X_i$ and $T_i$ may vary depending on the geophysical signal we
are trying to describe. To keep this section sufficiently general, we
hold off on specifying $X_i$ and $T_i$ until Section
\ref{sec:SignalModel}, where we estimate transient strain from slow
slip events in the Pacific Northwest.

GNSS data records transient displacements as well as other physical
and non-physical processes that we are not interested in. We first
consider component $i$ of a single GNSS displacement observation made
at station $j$, which is located at $\pos^{(j)}$, and time $t^{(k)}$.
We describe this observation, $d_i^{*(jk)}$, as a realization of the
random variable
\begin{align}\label{eq:Data}
\begin{split}
d_i^{(jk)} = &u_i\left(\pos^{(j)},t^{(k)}\right) + 
              \eta_i^{(jk)} + 
              a_i^{(j)} + b_i^{(j)}t^{(k)} + \\
             &c_i^{(j)}\sin\left(2 \pi t^{(k)}\right) +  
              e_i^{(j)}\cos\left(2 \pi t^{(k)}\right) + \\ 
             &f_i^{(j)}\sin\left(4 \pi t^{(k)}\right)  + 
              g_i^{(j)}\cos\left(4 \pi t^{(k)}\right), 
\end{split}
\end{align}
where $\eta_i^{(jk)}$ describes noise, $a_i^{(j)}$ is an offset that
is unique for each station, $b_i^{(j)}$ is the secular velocity at
$\pos^{(j)}$, and the sinusoids describe seasonal deformation (using
units of years for $t^{(k)}$). We then consider the column vector of
$n$ GNSS displacement observations, $\data^*_i$, where the subscript
indicates that we are still only considering component $i$ of
displacements. The observations are made at $m$ stations, and the
times and positions for each observation are described by the set
$\points$. $\data^*_i$ can be considered a realization of the random
vector $\data_i$, which is formed by evaluating eq. (\ref{eq:Data}) at
each point in $\points$. To write out $\data_i$ explicitly, we let
$\G$ be an $n \times 6m$ matrix consisting of the basis functions from
eq. (\ref{eq:Data}) (i.e., the linear trends and sinusoids for each
station) evaluated at each point in $\points$. The coefficients
corresponding to each basis function are collected into the column
vector $\mitbf{m}_i$. The noise for all the observations are described
by the column vector $\mitbf{\eta}_i$. We can then write $\data_i$ as
\begin{equation}\label{eq:DataVec}
\data_i = u_i(\points) + \mitbf{\eta}_i + \G\mitbf{m}_i,
\end{equation}
where the notation $u_i(\points)$ represents the column vector
$[u_i(p)]_{p \in \points}$.

We assume a diffuse prior for the components of $\mitbf{m}_i$, that is
to say $\mitbf{m}_i \sim \mathcal{N}(\zeros,\kappa^2\eye)$ in the
limit as $\kappa \to \infty$. Of course, the secular velocities,
$b_i^{(j)}$, are spatially correlated and we could invoke a tectonic
model to form a prior on $b_i^{(j)}$. However, in our application to
the Pacific Northwest, we will be using displacement time series which
are long enough to sufficiently constrain $b_i^{(j)}$ for each
station, avoiding the need to incorporate a prior. Likewise, seasonal
deformation is spatially correlated \citep{Dong2002,Langbein2008}, and
it may be worth exploring and exploiting such a correlation in a
future study. We assume that $\mitbf{\eta}_i$ is a spatially and/or
temporally correlated random vector distributed as
$\mathcal{N}(\zeros,\mitbf{C}_{\eta_i})$. For example,
$\mitbf{\eta}_i$ can be uncorrelated white noise, temporally
correlated noise describing benchmark wobble
\citep[e.g.,][]{Wyatt1982,Wyatt1989}, and/or spatially correlated
noise describing common mode error \citep[e.g.,][]{Wdowinski1997}. The
appropriate noise model may vary depending on the application, and we
hold off on specifying the covariance matrix, $\mitbf{C}_{\eta_i}$,
until Section \ref{sec:NoiseModel}. We are now able to write the
distribution of $\data_i$ as
\begin{equation}\label{eq:DataDist}
\data_i \sim \mathcal{N}\left(\zeros, C_{u_i}(\points,\points) + 
                                      \mitbf{C}_{\eta_i}  + 
                                      \kappa^2\G\G^T \right),
\end{equation}
where $C_{u_i}(\points,\points)$ represents the matrix
$[C_{u_i}(p,p')]_{(p,p') \in \points \times \points}$.

We form a posterior estimate of transient displacements, denoted as
$\post_i$, by updating $u_i$ with the fact that $\data_i^*$ was
realized from the random vector $\data_i$, that is to say $\post_i =
u_i |(\data_i = \data^*_i)$. A general solution for $\post_i$ is
derived in \citet[sec. 8.9]{VonMises1964}, where we find that
$\post_i$ is distributed as $\mathcal{GP}(\mu_{\post_i},C_{\post_i})$
with mean function
\begin{align}\label{eq:PosteriorMean}
\mu_{\post_i}(p) &= \E{u_i(p)} + 
                    \Cov{u_i(p),\data_i} 
                    \Cov{\data_i}^{-1}
                    \left(\data^*_i - \E{\data_i}\right)\nonumber \\
                 &= C_{u_i}(p,\points)\left( C_{u_i}(\points,\points) + 
                                             \mitbf{C}_{\eta_i} + 
                                             \kappa^2\G\G^T\right)^{-1}
                    \data^*_i
\end{align}    
and covariance function
\begin{align}\label{eq:PosteriorCov}
C_{\post_i}(p,p') &= \Cov{u_i(p),u_i(p')} - 
                     \Cov{u_i(p),\data_i} 
                     \Cov{\data_i}^{-1}
                     \Cov{\data_i,u_i(p')} \nonumber \\
                  &= C_{u_i}(p,p') - 
                     C_{u_i}(p,\points)\left( C_{u_i}(\points,\points) + 
                                              \mitbf{C}_{\eta_i} + 
                                              \kappa^2\G\G^T \right)^{-1}
                     C_{u_i}(\points,p').
\end{align}
However, we are interested in the limit as $\kappa \to \infty$, and
the form for eqs. (\ref{eq:PosteriorMean}) and (\ref{eq:PosteriorCov})
is not suitable for evaluating this limit. We use a partitioned matrix
inversion identity \citep[sec. 2.7.4]{Press2007} to rewrite eqs.
(\ref{eq:PosteriorMean}) and (\ref{eq:PosteriorCov}) as
\begin{equation}\label{eq:PosteriorMean2}
\mu_{\post_i}(p) = \left[\begin{array}{cc} 
                         C_{u_i}(p,\points) & \zeros \\
                         \end{array}\right]
                   \left[\begin{array}{cc}
                         C_{u_i}(\points,\points) + \mitbf{C}_{\eta_i} & \G \\
                         \G^T  & -\kappa^{-2} \eye \\
                         \end{array}\right]^{-1}
                   \left[\begin{array}{c}
                         \data^*_i \\
                         \zeros \\
                         \end{array}\right]
\end{equation}    
and
\begin{multline}\label{eq:PosteriorCov2}
C_{\post_i}(p,p') = C_{u_i}(p,p') - \\ 
                    \left[\begin{array}{cc}
                          C_{u_i}(p,\points) & \zeros \\
                          \end{array}\right]
                    \left[\begin{array}{cc}
                          C_{u_i}(\points,\points) + \mitbf{C}_{\eta_i} & \G \\
                          \G^T  & -\kappa^{-2} \eye \\
                          \end{array}\right]^{-1}
                    \left[\begin{array}{c}
                          C_{u_i}(\points,p') \\
                          \zeros \\
                          \end{array}\right].
\end{multline}
Taking the limit as $\kappa \to \infty$, we get the solution for the
mean and covariance of $\post_i$,
\begin{equation}\label{eq:PosteriorMean3}
\mu_{\post_i}(p) = \left[\begin{array}{cc}
                         C_{u_i}(p,\points) & \zeros \\
                         \end{array}\right]
                   \left[\begin{array}{cc}
                         C_{u_i}(\points,\points) + \mitbf{C}_{\eta_i} & \G \\
                         \G^T  & \zeros \\
                         \end{array}\right]^{-1}
                   \left[\begin{array}{c}
                         \data^*_i \\
                         \zeros \\
                         \end{array}\right]
\end{equation}    
and
\begin{equation}\label{eq:PosteriorCov3}
C_{\post_i}(p,p') = C_{u_i}(p,p') - 
                    \left[\begin{array}{cc}
                          C_{u_i}(p,\points) & \zeros \\
                          \end{array}\right]
                    \left[\begin{array}{cc}
                          C_{u_i}(\points,\points) + \mitbf{C}_{\eta_i} & \G \\
                          \G^T  & \zeros \\
                          \end{array}\right]^{-1}
                    \left[\begin{array}{c}
                          C_{u_i}(\points,p') \\
                          \zeros \\
                          \end{array}\right].
\end{equation}
To ensure that the inverse matrices in eqs. (\ref{eq:PosteriorMean3})
and (\ref{eq:PosteriorCov3}) exist, each column in $\G$ must be
linearly independent. This condition tends to be violated when there
are too few observations at a station. In that case, a singular value
decomposition can be used to remove linearly dependent components from
$\G$.

It should be noted that we have ignored any covariances between the
easting and northing components of $\vec{u}$ and $\vec{\data}$.
This simplification reduces the computational complexity of evaluating
the posterior transient displacements because each component can be
evaluated independently. However, we are inherently assuming that the
principle axes describing the distribution of $\vec{u}(p)$ and
$\vec{d}^{(jk)}$ are aligned with the cardinal directions, which is
admittedly an arbitrary assumption.

The posterior transient displacements are spatially and temporally
continuous, and we can use eqs. (\ref{eq:PosteriorMean3}) and
(\ref{eq:PosteriorCov3}) to evaluate $\post_i$ at any $p$.
Furthermore, $\post_i$ is spatially and temporally differentiable,
allowing us to formulate $\strain$ at any $p$ that we may be
interested in. The components of $\strain$ can be written as
\begin{equation}\label{eq:StrainRate}
\strain_{ij}(p) = \frac{1}{2} \frac{\partial}{\partial t} 
                  \left(\frac{\partial \post_i(p)}{\partial x_j} +  
                        \frac{\partial \post_j(p)}{\partial x_i} \right).
\end{equation}
Since eq. (\ref{eq:StrainRate}) is a linear operation on the Gaussian
processes $\post_i$ and $\post_j$, we know that $\strain_{ij}$ is also
a Gaussian process. From \citet[sec. 10]{Papoulis1991}, we find that
the mean and covariance functions for $\dot{\varepsilon}_{ij}$ are
\begin{align}\label{eq:StrainMean}
\mu_{\strain_\e\e}(p) &= \frac{\partial^2 \mu_{\post_\e}(p)}{\partial t \, \partial x_\e} \\
\mu_{\strain_\n\n}(p) &= \frac{\partial^2 \mu_{\post_\n}(p)}{\partial t \, \partial x_\n} \\
\mu_{\strain_\e\n}(p) &= \frac{1}{2}\frac{\partial}{\partial t}\left(
                                    \frac{\partial \mu_{\post_\e}(p)}{\partial x_\n} + 
                                    \frac{\partial \mu_{\post_\n}(p)}{\partial x_\e} \right)
\end{align} 
and
\begin{align}\label{eq:StrainCov}
C_{\strain_\e\e}(p,p') &= \frac{\partial^4 C_{\post_\e}(p,p')}{\partial t \, \partial t' \, \partial x_\e \, \partial x'_\e} \\
C_{\strain_\n\n}(p,p') &= \frac{\partial^4 C_{\post_\n}(p,p')}{\partial t \, \partial t' \, \partial x_\n \, \partial x'_\n} \\
C_{\strain_\e\n}(p,p') &= \frac{1}{4} \frac{\partial^2}{\partial t \, \partial t'}\left(
                                      \frac{\partial^2 C_{\post_\e}(p,p')}{\partial x_\n \, \partial x'_\n} + 
                                      \frac{\partial^2 C_{\post_\n}(p,p')}{\partial x_\e \, \partial x'_\e} \right),
\end{align} 
respectively.

%% TRANSIENT DETECTION CRITERION
\subsection{Transient detection criterion}\label{sec:TransientDetection}

Our motivation for estimating transient strain rates is, in part, to
detect geophysical phenomena. As we will see, geophysical signal can
be easily identified by visually inspecting the solution for $\strain$
from eqs. (\ref{eq:StrainMean}) and (\ref{eq:StrainCov}). However, if
we want to detect geophysical phenomena automatically, then we need to
define a detection criterion. We use a signal-to-noise ratio, SNR,
that is based on the Frobenius norm of $\strain$,
$||\strain||_\mathrm{F} = \left(\strain_{\e\e}^2 + \strain_{\n\n}^2 +
2\strain_{\e\n}^2\right)^{\frac{1}{2}}$, for our detection criterion.
In the geodetic literature, $||\dot{\varepsilon}||_\mathrm{F}$ is
often used as a metric for the strain rate ``magnitude", and it is
sometimes referred to as the second invariant of strain rate. Noting
that $||\strain||_\mathrm{F}$ is a random variable, we take SNR to be
the ratio of the estimated mean and standard deviation of
$||\strain||_\mathrm{F}$. An estimate of the mean is found by
evaluating $||\strain||_\mathrm{F}$ at the mean of $\strain$,
\begin{align}
\mu_{||\strain||_\mathrm{F}} &\approx ||\strain||_F \big|_{\strain = \mu_{\strain}}\nonumber \\
                             &= \left(\mu_{\strain_\e\e}^2 + 
                                      \mu_{\strain_\n\n}^2 + 
                                      2\mu_{\strain_\e\n}^2 \right)^\frac{1}{2},
\end{align}
and we use nonlinear uncertainty propagation to estimate the standard
deviation,
\begin{equation}
\sigma_{||\strain||_\mathrm{F}} \approx
\left( \left(\left. \frac{\partial ||\strain||_\mathrm{F}}{\partial \strain_\e\e}\right|_{\strain = \mu_{\strain}} \right)^2 
       \sigma^2_{\strain_\e\e} +
       \left(\left. \frac{\partial ||\strain||_\mathrm{F}}{\partial \strain_\n\n}\right|_{\strain = \mu_{\strain}} \right)^2
       \sigma^2_{\strain_\n\n} +
       \left(\left. \frac{\partial ||\strain||_\mathrm{F}}{\partial \strain_\e\n}\right|_{\strain = \mu_{\strain}} \right)^2 
       \sigma^2_{\strain_\e\n} \right)^\frac{1}{2},
\end{equation}
where $\sigma^2_{\strain_{ij}}(p) = C_{\strain_{ij}}(p,p)$. After some
calculations, we find SNR to be
\begin{align}\label{eq:SNR}
\mathrm{SNR}(p) &= \frac{\mu_{||\strain||_\mathrm{F}}(p)}{\sigma_{||\strain||_\mathrm{F}}(p)} \\
                &= \frac{\mu_{\strain_\e\e}(p)^2 +
                         \mu_{\strain_\n\n}(p)^2 +
                         2\mu_{\strain_\e\n}(p)^2}
                        {\big(\sigma^2_{\strain_\e\e}(p)\mu_{\strain_\e\e}(p)^2 + 
                              \sigma^2_{\strain_\n\n}(p)\mu_{\strain_\n\n}(p)^2 + 
                              4\sigma^2_{\strain_\e\n}(p)\mu_{\strain_\e\n}(p)^2
                         \big)^{\frac{1}{2}}}.
\end{align}
We explicitly show that SNR is a function of $p$ to emphasize that it
identifies the position and time of anomalous deformation. We can
reasonably suspect that some transient geophysical phenomena is
occurring wherever and whenever SNR is larger than ${\sim}3$.

%% OUTLIER DETECTION
\subsection{Outlier detection}\label{sec:Outlier}

In deriving our formulation for transient strain rates, we have
assumed that noise in the data vector is normally distributed. This is
not an appropriate assumption for GNSS data, which often have more
outliers than would be expected for normally distributed noise.
Methods for analyzing GNSS data should either be robust against
outliers or should involve a preprocessing step in which outliers are
detected and removed. Examples of the former include the MIDAS method
for estimating secular velocities \citep{Blewitt2016} and the GPS
Imaging method for detecting spatially coherent features
\citep{Hammond2016}. In this study, we identify and remove outliers as
a preprocessing step before estimating $\strain$. Outliers are
identified based on the residuals for a model that best fits the
observed data. Observations with residuals that exceed some threshold
are removed. This strategy for detecting outliers is commonly used for
GNSS data, where the model being fit to the data typically consists of
a linear trend and seasonal terms for each GNSS station
\citep[e.g.,][]{Johansson2002,Dong2006,Bos2013}. To prevent transient
geophysical signal from being erroneously identified as outliers, the
model used in our outlier detection algorithm additionally consists of
a temporally correlated Gaussian process. The details of our algorithm
are given in Appendix A.

It should be noted that our algorithm does not identify jumps in GNSS
time series, which are another common issue. Some, but not all, jumps
can be automatically removed by looking up the dates of equipment
changes and earthquakes \citep{Gazeaux2013}. However, it is still
necessary to manually find and remove jumps of unknown origin.

%% Application to Pacific Northwest Slow Slip Events
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Application to Pacific Northwest Slow Slip Events}\label{sec:Cascadia}

We use our method to estimate transient strain rates in the Pacific
Northwest, and we are specifically interested in identifying strain
resulting from SSEs \citep[e.g.,][]{Dragert2001}. In the Pacific
Northwest, SSEs can be detected by monitoring for associated seismic
tremor \citep{Rogers2003}, which is actively being done by the Pacific
Northwest Seismic Network \citep{Wech2010}. We can thus compare the
tremor records to the transient strain rates estimated with GPR to
verify that we are indeed identifying strain from SSEs.

We use the daily displacement solutions generated by the Geodesy
Advancing Geosciences and EarthScope (GAGE) facility for continuous
GNSS stations \citep{Herring2016}. This data is publicly available and
can be found at www.unavco.org. We limit the dataset to the stations
and time ranges which are pertinent to the seven most recent SSEs in
the Puget Sound region. The earliest SSE considered in this study
began in August 2010, and the most recent SSE began in February 2017.
We use these most recent SSEs because the station coverage is
sufficiently dense for us to use maximum likelihood methods to
constrain prior models.  The positions of GNSS stations used to
estimate transient strain rates are shown in Figure \ref{fig:Context}.

\begin{figure*}
\includegraphics{figures/context_map/context-map.pdf}
\caption{
Positions of continuous GNSS stations used to estimate transient
strain rates. The colored regions indicate the distribution of seismic
tremor as determined by \citet{Wech2010}. The red dots show the
positions of GNSS stations mentioned in this paper. The blue dot
indicates the location of the transient strain rates shown in Figure
\ref{fig:StrainTs} and the signal-to-noise ratio shown in Figure
\ref{fig:StrainMag}. The blue dashed circle demarcates the spatial
extent of the tremors shown in Figure \ref{fig:StrainMag}.
}    
\label{fig:Context}
\end{figure*}

%% NOISE MODEL
\subsection{Noise model}\label{sec:NoiseModel}

Before we determine the transient strain rates, we must establish a
prior for the transient displacements, $u$, and the noise, $\eta$. In
this section we discuss our choice for the noise covariance function
$C_\eta$. There have been numerous studies on temporally correlated
noise in GNSS data \citep[e.g.,][]{Zhang1997, Mao1999, Williams2004,
Langbein2008}. In these studies, temporally correlated noise was
described with some combination of Brownian motion, a first-order
Gauss-Markov (FOGM) process, and/or flicker noise. There is some
physical justification for using Brownian motion as a noise model
because it accurately describes the power spectrum of motion resulting
from instability in geodetic monuments \citep[e.g.,][]{Wyatt1982,
Wyatt1989}. Here we describe the time dependence of $\eta$ as a FOGM
process and consider $\eta$ to be spatially uncorrelated. A FOGM
process is a solution to the stochastic differential equation
\begin{equation}\label{eq:FOGMdef}
\dot{\eta}(t) + \alpha \eta(t) = \beta w(t),
\end{equation}
where $w(t)$ is white noise with unit variance. The FOGM process
degenerates to the commonly used Brownian motion noise model under the
condition that $\alpha=0$ and $\eta(0) = 0$. Our spatially
uncorrelated noise model that satisfies eq. (\ref{eq:FOGMdef}) is a
Gaussian process with zero mean and the covariance function
\begin{equation}\label{eq:FOGM}
C_\eta\left((\pos,t),(\pos',t')\right) = 
\begin{cases}
\frac{\beta^2}{2\alpha}\exp\left(-\alpha|t - t'|\right), &\pos = \pos' \\
0, &\pos \neq \pos'
\end{cases}. 
\end{equation}

We constrain the hyperparameters for $\eta$, $\alpha$ and $\beta$,
with a set of 38 continuous GNSS stations in the Pacific Northwest
that are east of 121$^\circ$W.  These stations are sufficiently far
from the subduction zone that they are unlikely to contain transient
signal associated with SSEs.  We clean the data for these stations by
removing jumps at times of equipment changes, and we remove outliers
that have been detected with the algorithm described in Section
\ref{sec:Outlier}. We then find $\alpha$ and $\beta$ for each station
time series with the Restricted Maximum Likelihood (REML) method
\cite[e.g.,][]{Harville1974,Cressie1992,Hines2017}. The REML method
finds the hyperparameters, which we collectively refer to as
$\mitbf{\theta}$, that maximize the likelihood function
\begin{equation}\label{eq:REML}
\mathcal{L}(\mitbf{\theta}) = \left(\frac{\left|\G^T\G\right|}{(2\pi)^{n-6m} 
                                    \left| \mitbf{\Sigma}(\mitbf{\theta}) \right| 
                                    \left| \G^T\mitbf{\Sigma}(\mitbf{\theta})^{-1}\G \right|}\right)^{\frac{1}{2}} 
                              e^{-\tfrac{1}{2}\data_*^T\mitbf{K}(\mitbf{\theta})\data_*},
\end{equation}
where
\begin{equation}
\mitbf{K}(\mitbf{\theta}) = \mitbf{\Sigma}(\mitbf{\theta})^{-1} - 
                            \mitbf{\Sigma}(\mitbf{\theta})^{-1}\G
                            \left(\G^T\mitbf{\Sigma}(\mitbf{\theta})^{-1}\G\right)^{-1}
                            \G^T\mitbf{\Sigma}(\mitbf{\theta})^{-1}.
\end{equation}
\citet{Harville1974} showed that choosing the hyperparameters which
maximize eq. (\ref{eq:REML}) is equivalent to choosing the
hyperparameters which maximize the probability of drawing $\data_*$
from $\data$. We use the REML method over the maximum likelihood
method \citep[e.g.,][]{Langbein1997} because the REML method accounts
for the improper prior that we assigned to $\mitbf{a}$
\citep{Hines2017}. We independently estimate $\mitbf{\theta}$ for each
station, and so $\data_*$ consists of displacements for an individual
station. We are assuming $u(p)=0$ when estimating the noise
hyperparameters for this section. The distribution of inferred
$\alpha$ and $\beta$ are shown in Figure \ref{fig:NoiseParams}. The
amplitude of FOGM noise, $\beta$, for the easting and northing
components is notably low and are clustered around 0.5 mm/yr$^{0.5}$.
The corresponding estimates of $\alpha$ tend to cluster around 0
yr$^{-1}$, suggesting that noise can be described with Brownian
motion. We also estimate hyperparameters for the vertical component of
displacements, under the hope that vertical deformation gradients
could reveal some geophysical signal. The amplitude of FOGM noise for
the vertical component is relatively large with a median value of 13.5
mm/yr$^{0.5}$.  The inferred values for $\alpha$ are also higher for
the vertical component with a median value of 8.21 yr$^{-1}$. In
Figure \ref{fig:NoiseSamples}, we use the median values of $\alpha$
and $\beta$ to generate two random samples of FOGM noise for each
component. The samples span five years and over these five years the
easting and northing samples drift by about 1 mm. In the context of
detecting SSEs, which produce several mm's of surface displacement on
the time-scale of weeks, the estimated FOGM noise for the easting and
northing component is negligible. In contrast, the estimated FOGM
noise for the vertical component is larger than the signal we would
expect from SSEs. We suspect that the higher amplitude for the FOGM
noise in the vertical component is accommodating for deficiencies in
our rather simple seasonal model. Based on this analysis, we
henceforth ignore temporally correlated noise in the easting and
northing component because of its low amplitude. We also do not use
vertical displacements because of the presumably low signal-to-noise
ratio.

\begin{figure*}
\includegraphics{figures/noise/noise-params.pdf}
\caption{
Distribution of estimated FOGM hyperparameters (eq. \ref{eq:FOGM}).
Hyperparameters are estimated with the REML method for 38 stations in
the Pacific Northwest that are east of $121^\circ$W. ``IQR'' is the
interquartile range.
}   
\label{fig:NoiseParams}
\end{figure*}

\begin{figure*}
\includegraphics{figures/noise/noise-samples.pdf}
\caption{
Two FOGM noise samples for each component. The FOGM hyperparameters
have been set to the median values from Figure \ref{fig:NoiseParams}.
}   
\label{fig:NoiseSamples}
\end{figure*}

Another significant source of noise in GNSS data is common mode error
\citep[e.g.,][]{Wdowinski1997,Dong2006}, which is noise that is highly
spatially correlated. When not accounted for, common mode error
manifests as spatially uniform undulations in estimated transient
displacements. However, estimated transient strain rates are
insensitive to common mode error. We therefore do not include common
mode error in our noise model. We then make the simplifying assumption
that $\eta(p) = 0$ for the easting and northing component of GNSS
data.

%% PRIOR MODEL
\subsection{Prior model}\label{sec:SignalModel}

We next establish our prior model for transient displacements.
Specifically, we discuss our choice for the covariance functions
$X(\pos,\pos')$ and $T(t,t')$. For $X$, we use the squared exponential
(SE) covariance function,
\begin{equation}\label{eq:SE}
X(\pos,\pos') = \exp\left(\frac{-||\pos - \pos'||_2^2}{2 \ell^2}\right).
\end{equation}
The SE covariance function is commonly used in kriging
\citep[e.g,][]{Cressie1992} and Gaussian process regression
\citep[e.g.,][]{Rasmussen2006}. In terms of geodetic applications,
\citet{Kato1998} and \cite{El-Fiky1999} demonstrated that the SE
accurately describes the covariance of secular GNSS derived velocities
in Japan. The SE is a positive definite covariance function for any
number of spatial dimensions. A Gaussian process with an SE covariance
function is isotropic and has realizations that are infinitely
differentiable.

We consider three potential models for the temporal covariance of $u$.
First, we consider the one-dimensional SE covariance function,
\begin{equation}\label{eq:TimeSE}
T(t,t') = \phi^2\exp\left(\frac{-|t - t'|^2}{2\tau^2}\right).
\end{equation}
Note that $T$ includes the hyperparameter $\phi$, which serves to
scale the covariance function $C_u$. Second, we consider integrated
Brownian motion (IBM). IBM has zero mean and its covariance function
can be found by integrating the covariance function for Brownian
motion as
\begin{align}\label{eq:IBM}
T(t,t') &= \int_0^t \int_0^{t'} \phi^2 \min(s,s') \,ds'\,ds \\
        &= \frac{\phi^2}{2}\min(t,t')^2 \left(\max(t,t') - \frac{1}{3}\min(t,t')\right), \ \ \ t,t' \geq 0.
\end{align}
IBM has been used in the context of Kalman filtering as a
non-parametric model for the time dependence of geophysical signals
\citep[e.g.,][]{Segall1997, McGuire2003, Ohtani2010, Hines2016a}. It
should be emphasized $t=0$ is a reference time at which the Gaussian
process is exactly zero. For some geophysical signals, it is
appropriate to have this reference time. For example, if we are trying
to identify postseismic deformation then $t$ should be zero at the
time of the earthquake.  However, if we are interesting in detecting
transient events, where there is no known start time, then IBM may not
be an appropriate prior, and an isotropic Gaussian process should be
preferred. In the following analysis, we make the quite arbitrary
choice that $t$ is zero on the first epoch of $\data_*$. Using an
earlier reference time does not change the results discussed in this
section. Our third option for $T$ is the Wendland class of covariance
functions \citep{Wendland2005}. Wendland covariance functions have
compact support and hence their corresponding covariance matrices are
sparse. In our analysis, we exploit this sparsity with the CHOLMOD
software package \citep{Chen2008}. Wendland functions are positive
definite in $\mathbb{R}^d$, and they describes an isotropic Gaussian
process with realizations that can be differentiated $k$ times. The
form of the covariance function depends on the choice of $d$ and $k$.
We use $d=1$ since we are describing the temporal covariance of $u$.
We use $k=2$, giving samples of $u$ continuous velocities and
accelerations. The corresponding Wendland covariance function is
\begin{equation}\label{eq:Wendland}
T(t,t') = \phi^2\left(1 - \frac{|t - t'|}{\tau}\right)^5_+ \left(\frac{8|t - t'|^2}{\tau^2} + \frac{5|t - t'|}{\tau} + 1\right), 
\end{equation}
where
\begin{equation}
(t)_+ = 
\begin{cases}
t, \ \ \ t > 0 \\
0, \ \ \ \mathrm{otherwise}.
\end{cases}
\end{equation}

We next determine appropriate hyperparameters for $X$ and each of the
three candidate covariance functions for $T$. First, we clean the GNSS
datasets by removing offsets at times of equipment changes and
removing outliers with the method describe in Section
\ref{sec:Outlier}. For the outlier detection algorithm, our prior
model, $u$, is chosen to have a length-scale and time-scale that is
able to approximately describe SSE displacements. We use the SE
covariance function for $X$ with length-scale $\ell = 100$ km, and we
use the Wendland covariance function for $T$, due to its computational
efficiency, with time-scale $\tau = 0.1$ yr and $\phi = 1$ mm.  The
outlier detection algorithm is particularly effective at removing
outliers for stations at high elevation (Figure \ref{fig:Outliers}),
which can be adversely affected by ice or snow during the winter
\citep{Lisowski2008}. After cleaning the dataset, we divide it into
seven subsets which are four months long and each centered on the time
of a SSE. The times of the seven SSEs are determined with tremor
records from \cite{Wech2010}. We use the REML method to find the
optimal hyperparameters for $T$ and $X$ for each subset of data. We
choose to make each data subsets four months long because it is long
enough to encompass a SSE in the Pacific Northwest, while it is short
enough to still be computationally tractable. However, four months is
too short to resolve the sinusoids in $\data$, and they are omitted
from $\data$ in this REML analysis for Pacific Northwest SSEs. The
estimated hyperparameters for $u$ are summarized in Table 1. Based on
the interquartile ranges, the estimated hyperparameters for the SE and
Wendland covariance functions do not vary significantly between SSEs.
This suggests that the median of estimated hyperparameters should be
an appropriate prior model for all Pacific Northwest SSEs. For the IBM
model, there are several anomalously large values for $\ell$ and
$\phi$, which results in large interquartile ranges.

\begin{figure*}
\includegraphics{figures/outliers/outliers.pdf}
\caption{
Detrended easting component of displacements at station SC03, which is
located on Mount Olympus in Washington. The orange markers indicate
outliers that were automatically detected using the algorithm from
Section \ref{sec:Outlier}. The error bars show one standard deviation
uncertainties. Note that outliers tend be observed in the winter,
suggesting that they were caused by snow or ice.
}   
\label{fig:Outliers}
\end{figure*}

\begin{table}\label{tab:Parameters}
\caption{
Optimal hyperparameters for the prior on transient displacements
determined with the REML method. The temporal covariance function is
indicated by the ``$T$'' column. The SE, IBM, and Wendland covariance
functions are defined in eqs. (\ref{eq:TimeSE}), (\ref{eq:IBM}), and
(\ref{eq:Wendland}), respectively. The spatial covariance function,
$X$, is the squared exponential (eq. \ref{eq:SE}) in all cases. The
hyperparameters are estimated for each of the seven SSEs considered in
this study, and the tabulated values indicate the median and
interquartile ranges of estimates. The ``diff $\log$(REML)'' column
compares the log REML likelihood to the log REML likelihood when using
the SE covariance function for $T$. Negative values indicate that
observations are more consistent with the SE covariance function.
} 
\begin{tabular} {l l l l l l}
$T$ & direction & $\ell$  & $\phi$   & $\tau$  & diff. $\log$(REML) \\ \hline
SE & east   & 92 $\pm$ 25 km  & 0.62 $\pm$ 0.11 mm  & 0.026 $\pm$ 0.011 yr  &  - \\
SE & north  & 91 $\pm$ 53 km  & 0.43 $\pm$ 0.05 mm  & 0.030 $\pm$ 0.017 yr  &  - \\
Wendland & east   & 95 $\pm$ 30 km  & 0.66 $\pm$ 0.15 mm  & 0.093 $\pm$ 0.044 yr &  0.78 $\pm$ 0.87 \\
Wendland & north  & 92 $\pm$ 57 km  & 0.46 $\pm$ 0.10 mm  & 0.116 $\pm$ 0.057 yr &  0.08 $\pm$ 0.58 \\
IBM & east   & 110 $\pm$ 130 km & 290 $\pm$ 420 mm/yr$^{1.5}$  & -          & -16.4 $\pm$ 7.8 \\
IBM & north  & 150 $\pm$ 560 km & 110 $\pm$ 250 mm/yr$^{1.5}$ & -           & -10.1 $\pm$ 2.3 \\
\end{tabular}
\end{table}

Next we identify which covariance function for $T$ best describes the
SSEs. One approach is to compare the REML likelihoods for each
covariance function, similar to the analysis in \citet{Langbein2004}.
In Table 1, we summarize how the log REML likelihoods for the Wendland
and IBM covariance functions compare to the SE covariance function.
Based on the differences in log REML likelihoods, the data is
substantially more likely to come from a Gaussian process with a SE or
Wendland covariance function than an IBM covariance function. The REML
likelihoods do not definitively indicate whether the SE or Wendland
covariance function is preferable.

To further explore which covariance function for $T$ best describes
SSEs, we compare the observations to the predicted displacements for
each covariance function. We consider the data prediction vector to be
$\hat{\data} = \left(u(\points) + \G\mitbf{a}\right)|\data_*$.
Following a similar procedure as in Section \ref{sec:Method}, it can
be shown that $\hat{\data}$ is normally distributed with mean
\begin{equation}\label{eq:DataPredMean}
\mitbf{\mu}_{\hat{d}} = \left[\begin{array}{cc}
                           C_u(\points,\points) & \G \\
                           \end{array}\right]
                     \left[\begin{array}{cc}
                           \mitbf{\Sigma} & \G \\
                           \G^T  & \zeros \\
                           \end{array}\right]^{-1}
                     \left[\begin{array}{c}
                           \data_* \\
                           \zeros \\
                           \end{array}\right]
\end{equation}  
and covariance
\begin{equation}\label{eq:DataPredCov}
\mitbf{C}_{\hat{\data}} = C_u(\points,\points) - 
                        \left[\begin{array}{cc}
                              C_u(\points,\points) & \G \\
                              \end{array}\right]
                        \left[\begin{array}{cc}
                              \mitbf{\Sigma} & \G \\
                              \G^T  & \zeros \\
                              \end{array}\right]^{-1}
                        \left[\begin{array}{c}
                              C_u(\points,\points) \\
                              \G^T \\
                              \end{array}\right].
\end{equation}
We compute $\hat{\data}$ using SE, Wendland, and IBM covariance
functions for $T$ and the median hyperparameters from Table 1. Figure
\ref{fig:Fit} compares the easting component of $\data_*$ to
$\hat{\data}$ during the winter 2015-2016 SSE at the three stations
that record the strongest SSE signal, ALBH, P436, and SC02. Based on
Figure \ref{fig:Fit}, $\hat{\data}$ is insensitive to the choice of
covariance function for $T$. The predicted displacements for the IBM
covariance function contain slightly more high frequency, and perhaps
spurious, features. When comparing $\hat{\data}$ to $\data_*$ at
station ALBH and SC02, it appears that $\hat{\data}$ slightly
underestimates the rate of deformation during the SSE, regardless of
the chosen covariance function. This over-smoothing could indicate
that the chosen time-scale hyperparameter, $\tau$, is too large. The
predicted displacements at the remaining stations, which record more
subtle SSE deformation, seem to faithfully describe $\data_*$.

For our estimates of transient strain discussed in the next section,
we ultimately settle on the Wendland covariance function for $T$ and
use the median values from Table 1 for the hyperparameters. We choose
the Wendland covariance function over the SE covariance function
because of its computational advantages.

\begin{figure*}
\includegraphics{figures/signal_fit/signal-fit.pdf}
\caption{
Easting component of the observed displacements, $\data_*$, and
predicted displacements, $\hat{\data}$, during an SSE at three
stations where the SSE signal is strongest. The predicted
displacements are shown for when $T$ is a squared exponential (SE),
Wendland, and integrated Brownian motion (IBM) covariance function.
The one standard deviation uncertainties are shown for the
observations and the SE predictions. For clarity, uncertainties are
not shown for the IBM and Wendland predictions, but they are nearly
equivalent to the uncertainties for the SE predictions. The SE and
Wendland predictions are practically indistinguishable.
}   
\label{fig:Fit}
\end{figure*}

%% TRANSIENT STRAIN RATES
\subsection{Transient Strain Rates}\label{sec:Results}

Having established a noise model and a prior for transient
displacements, we use the cleaned GNSS dataset to calculate transient
strain rates in the Puget Sound region.  We calculate transient strain
rates for each day from January 1, 2010 to May 15, 2017. The strain
rates are estimates at a grid of points spanning the study area. In
Figure \ref{fig:StrainMap} we show the transient strain rates on
January 1, 2016, which coincides with the height of the winter
2015-2016 SSE. We have included an animation showing the map view of
strain rates through time as supplementary material. The strain rates
shown in Figure \ref{fig:StrainMap} are generally similar to the
strain rates during the other six SSEs considered in this study. The
SSEs cause trench perpendicular compression in the Olympic Peninsula
and extension east of Puget Sound. The strain transitions from
compression to extension around the southern tip of Vancouver Island,
which coincides with the location of thrust slip for SSEs in the Puget
Sound region \citep[e.g.,][]{Dragert2001,Wech2009,Schmidt2010}. Thus,
this pattern of strain is to be expected. During the period in between
SSEs, secular strain rates indicate trench perpendicular compression
throughout this study region \citep{Murray2000, McCaffrey2007,
McCaffrey2013}. When comparing inferred strain rates from SSEs to the
secular strain rates, we see that SSEs are concentrating tectonically
accumulated strain energy towards the trench, and presumably pushing
the subduction zone closer to failure.

%A key difference between the strain inferred here and strain that can
%be derived from fault slip models is that our estimated strain rates
%are not based on an assumed physical model. In contrast, fault slip
%models can be biased by errors in the assumed fault geometry or
%lithospheric rheology. Moreover, the degrees of freedom in fault slip
%models usually cannot be constrained by GNSS data alone, and it is
%necessary to impose regularization which further biases the results.
%Since our estimated strain rates lack such systematic errors, we can
%be more confident that our solution is unbiased and has meaningful
%uncertainties.

\begin{figure*}
\includegraphics{figures/strain_map/strain-map.pdf}
\caption{
Estimated transient strain rates during the Winter 2015-2016 SSE.
Strain glyphs show the normal strain rate along each azimuth, where
orange indicates compression and blue indicates extension. The shaded
regions indicate one standard deviation uncertainties in the normal
strain rates.
}   
\label{fig:StrainMap}
\end{figure*}

In Figure \ref{fig:StrainTs} we show the time dependence of estimated
transient strain rates at a position on the Olympic Peninsula, where
transient strain rates from SSEs are largest. To verify that the
estimated transient strain rates are accurately identifying
geophysical signal, we compare the signal-to-noise ratio from eq.
(\ref{eq:SNR}) to the frequency of seismic tremor (Figure
\ref{fig:StrainMag}). A signal-to-noise ratio greater than ${\sim}3$
can be interpreted as a detected geophysical signal. We detect nine
distinct events, which each correspond to peaks in seismic tremor. The
smaller events detected in August 2014 and February 2017 can be
considered inter-SSE events. They were not among the SSEs used to
constrain the prior covariance function. In between peaks in seismic
tremor, the signal-to-noise ratio is consistently between 0 and 2,
suggesting that all the transient strain detected at this location is
associated with SSEs and inter-SSE events.

\begin{figure*}
\includegraphics{figures/strain_ts/strain-ts.pdf}
\caption{
Three components of the transient horizontal strain rate tensor
estimated at the position shown in Figure \ref{fig:Context}. The
shaded regions indicate one standard deviation uncertainty.
}   
\label{fig:StrainTs}
\end{figure*}

\begin{figure*}
\includegraphics{figures/strain_ts/mag-ts.pdf}
\caption{
(top) Signal-to-noise ratio (eq. \ref{eq:SNR}) at the position shown
in Figure \ref{fig:Context}. (bottom) Frequency of tremors in the
region shown in Figure \ref{fig:Context}.
}   
\label{fig:StrainMag}
\end{figure*}

The results we have presented thus far indicate that we are
identifying the strain that we should expect to see. There are,
however, subtle features in our estimated transient strain rates which
we were not expecting. For example, some SSEs are preceded by a brief
period of east-west extension on the Olympic Peninsula. This feature
can be seen several days before the summer 2012 and winter 2015-2016
SSEs in Figure \ref{fig:StrainTs} as well as in the supplementary
animation. While this deformation is noteworthy, a discussion on the
mechanisms causing it is outside the scope of this study.

%% DISCUSSION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}\label{sec:Discussion}
Our results demonstrate that GPR is an effective tool for estimating
transient strain from GNSS data, which can then be used to detect
geophysical processes. One may argue that geophysical signal can also
be detected by merely inspecting the GNSS displacement time series.
Indeed, the SSEs identified in Figure \ref{fig:StrainMag} do produce
visible displacements in the GNSS data. However, the GNSS data also
contains outliers and non-tectonic deformation that is localized to
individual stations. In contrast, our estimates of transient strain
only identify features that are sufficiently spatially and temporally
coherent, based on our chosen prior model. Furthermore, our estimates
of transient strain are insensitive to common mode noise, which is
highly spatially correlated noise resulting from factors such as
reference frame error. Common mode noise can obscure geophysical
signal in the GNSS data, but it gets canceled out when computing the
transient strain. Lastly, our estimates of transient strain rates are
a spatial and temporal derivative of displacements, and thus any
geophysical signal in the transient strain rates tends to be more
pronounced than in the GNSS data. For these reasons, we argue that
transient strain rates estimated with the method described in Section
\ref{sec:Method} can illuminate geophysical signal that may not be
discernible from the noise in the GNSS displacement data.

In addition to detecting geophysical processes, the GNSS derived
transient strain rates can be used to better understand the data from
borehole strain meters (BSMs). The Plate Boundary Observatory
maintains about forty BSMs in the Pacific Northwest, and it has been
demonstrated that BSMs are able to record transient geophysical events
such as SSEs \citep[e.g.,][]{Dragert2011}. However, there are
complications that prevent BSM data from being used quantitatively in
geophysical studies. One difficulty is that BSM data should be
calibrated with a well known strain source, such as diurnal and
semidiurnal tides \citep{Hart1996,Roeloffs2010,Hodgkinson2013}.
Unfortunately, the tidal forces at BSMs which record SSEs are strongly
influenced by local bodies of water such as the Straight of Juan de
Fuca, making it difficult to form a theoretical prediction of tidal
strains \citep{Roeloffs2010}. Another complication is that noise in
BSM data is not well understood. The noise consists, in part, of a
long-term decay resulting from the instrument equilibrating with the
surrounding rock \citep{Gladwin1987}. Typically, this noise is dealt
with in an ad-hoc manner by fitting and removing exponentials and
low-order polynomials. We envision that the GNSS derived strain rates
from this paper can be used as a reference strain for calibrating BSM
data and quantify its noise.

There is potential for a more thorough analysis of the spatio-temporal
noise in GNSS data, $\eta$, than what was performed in Section
\ref{sec:NoiseModel}. We did not explore the spatial covariance of
$\eta$, which would describe common mode noise. We are able to ignore
common mode error in this study; however, for other geophysical
studies based on GNSS data, such as fault slip inversions, it may be
necessary to incorporate a spatially covarying noise model
\citep[e.g.,][]{Miyazaki2003}. We can also improve upon the seasonal
model used in this study, which consists of four spatially
uncorrelated sinusoids for each station. We did not explore the
spatial covariance of seasonal deformation or the temporal roughness
(i.e., the number of sinusoids needed to describe the observations).
The periodic Gaussian process \citep{Mackay1998} is an alternative
model for seasonal deformation and is well suited for exploring the
roughness of seasonal deformation.  The periodic Gaussian process has
zero mean and the covariance function
\begin{equation}\label{eq:Periodic}
T(t,t') = \phi^2 \exp\left(\frac{-\sin(\pi|t - t'|)^2}{2\tau^2}\right).
\end{equation}
Realizations have annual periodicity and the roughness is controlled
by $\tau$. Decreasing $\tau$ has the same effect as including higher
frequency sinusoids in the seasonal model. The optimal value for
$\tau$ can be found with the REML method as described in Section
\ref{sec:NoiseModel}.

The transient strain rates estimated in this study are constrained by
about seven years of daily displacement observations from 94 GNSS
stations. It can be computationally intensive to evaluate eqs.
(\ref{eq:PosteriorMean3}) and (\ref{eq:PosteriorCov3}) for a dataset
with this size. We significantly reduce the amount of memory needed to
estimate transient strain rates by describing the temporal covariance
of displacements with a compact Wendland covariance function. Using a
compact covariance function for our prior turns eqs.
(\ref{eq:PosteriorMean3}) and (\ref{eq:PosteriorCov3}) into sparse
systems of equations, which we then solve with CHOLMOD. CHOLMOD is
designed for solving sparse, positive definite systems of equations.
The matrix being inverted in eqs. (\ref{eq:PosteriorMean3}) and
(\ref{eq:PosteriorCov3}) is not positive definite; however, we can use
another partitioned matrix inversion identity from \citet{Press2007}
to partition it into positive definite submatrices to be inverted.
Even when using a compact covariance function, it may still be
necessary to reduce the computational burden by dividing the data into
subsets and evaluating transient strain rates for each subset.

%% CONCLUSION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}\label{sec:Conclusion}
In this paper we propose using Gaussian process regression (GPR) to
estimate transient strain rates from GNSS data. Most other methods for
estimating strain rates assume a parametric representation of
deformation, which can bias the results if the parameterization is not
chosen carefully. Here we assume a stochastic, rather than parametric,
prior model for displacements. Our prior model describes how much we
expect transient displacements to covary spatially and temporally. If
we know nothing about the underlying signal that we are trying to
recover, then the prior model can be chosen objectively with maximum
likelihood methods. Because GPR is a Bayesian method, the
uncertainties on our estimated transient strain rates are well
quantified, allowing one to discern geophysical signal from noise. We
demonstrate that GPR is an effective tool for detecting geophysical
phenomena, such as slow slip events, in our application to GNSS data
from the Pacific Northwest. One limitation with GPR is that it is not
robust against outliers. To overcome this limitation, we have
introduced an effective pre-processing method for identifying and
removing outliers from GNSS datasets. Another complication with GPR is
that it usually involves inverting a dense matrix where the number of
rows and columns is equal to the number of observations. This is
prohibitive when using several years of daily GNSS observations from a
network of several hundred stations. We significantly reduce the
computational burden of GPR by using compact Wendland covariance
function to describe our prior model. While this paper just focuses on
estimating transient strain rates, we believe that GPR is a powerful
tool that can be applied to a wide range of geophysical problems.

%% ACKNOWLEDGEMENTS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Acknowledgements}
This material is based upon work supported by the National Science
Foundation under grant EAR 1245263. The EarthScope Plate Boundary
Observatory data is provided by UNAVCO through the GAGE Facility with
support from the National Science Foundation (NSF) and National
Aeronautics and Space Administration (NASA) under NSF Cooperative
Agreement EAR-1261833. An implementation of the method described in
this paper is named Python-based Geodetic Network Strain software
(PyGeoNS). PyGeoNS is distributed under the MIT License and can be
found at www.github.com/treverhines/PyGeoNS.

%% APPENDIX
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix
\section{Outlier detection algorithm}

Our outlier detection algorithm is loosely based on the data editing
algorithm from \citet{Acheson1975}. Let $\data^*$ denote all $n$ GNSS
displacement observations for a single directional component, which
have been made at positions and times $\points$.  We describe
$\data^*$ as a realization of the random vector
\begin{equation}\label{eq:OutlierDataVec}
\data = \G\mitbf{m} + v(\points) + \mitbf{w},
\end{equation}
where $\G$ and $\mitbf{m}$ are the same as in eq. (\ref{eq:DataVec}),
$v$ is a Gaussian process distributed as $\mathcal{GP}(0,C_v)$, and
$\mitbf{w}$ is a vector of uncorrelated Gaussian noise with known
standard deviations $\mitbf{\sigma} = [\sigma_1, \sigma_2, ...,
\sigma_n]$. The Gaussian process $v$ is intended to describe transient
features in the data that cannot be explained by the linear trend or
seasonal terms in $\G$. We let the temporal covariance of $v$ be a
squared exponential, and we let $v$ be spatially uncorrelated so that,
\begin{equation}
C_v(p,p') = \phi^2\exp\left(\frac{-|t - t'|^2}{2\tau^2}\right) \delta_{\pos,\pos'},
\end{equation}
where $\delta_{\pos,\pos'} = 1$ if $\pos = \pos'$ and $0$ otherwise.
The spatial covariance of $v$ has little effect on the detected
outliers, and so we have assumed that $v$ is spatially uncorrelated
for simplicity. Based on our experience, $v$ can reasonably describe
most transient features in the data when we set $\phi = 1$ mm and
$\tau = 10$ days.

Our goal is to find the index set of non-outliers in $\data^*$, which
we denote as $\mitbf{\Omega}$. We use a tilde to indicate that an
array only contains elements corresponding to $\mitbf{\Omega}$ (e.g.,
the vector of non-outlier observations is denoted $\tilde{\data}^* =
[d_i^*]_{i \in \mitbf{\Omega}}$). The outliers are identified
iteratively, and we initiate $\mitbf{\Omega}$ with all $n$ indices. We
consider outliers to be data that are poorly explained by the model
$\G\mitbf{m} + v(\points)$, which is determine by the residual vector
\begin{align}\label{eq:Residual} 
\mitbf{r} &= \data^* - \E{\Big(\G\mitbf{m} + v(\points)\Big) \Big| \Big(\tilde{\data} =
                                                                        \tilde{\data}^*\Big)} \\
          &= \data^*  - 
            \left[\begin{array}{cc}
                  C_v(\points,\tilde{\points}) & \G \nonumber \\
                  \end{array}\right]
            \left[\begin{array}{cc}
                  C_v(\tilde{\points},\tilde{\points}) + \mathrm{diag}(\tilde{\mitbf{\sigma}}^2) & \tilde{\G} \\
                  \tilde{\G}^T  & \zeros \\
                  \end{array}\right]^{-1}
            \left[\begin{array}{c}
                  \tilde{\data}^* \\
                  \zeros \\
                  \end{array}\right].
\end{align}
Data with abnormally large residuals are identified as outliers. For
each iteration, we compute $\mitbf{r}$ and then update
$\mitbf{\Omega}$ so that it contains the indices of $\mitbf{r}$ whose
weighted values are less than $\lambda$ times the weighted root mean
square of $\tilde{\mitbf{r}}$,
\begin{equation}\label{eq:Update}
\mitbf{\Omega} \leftarrow \left\{i : \left|\frac{r_i}{\sigma_i}\right| < \lambda \cdot \sqrt{\frac{1}{|\mitbf{\Omega}|} \sum_{j \in \mitbf{\Omega}} \frac{r_j^2}{\sigma_j^2}} \right\}.
\end{equation} 
Iterations continue until the new $\mitbf{\Omega}$ is the same as the
previous $\mitbf{\Omega}$.

The outlier detection algorithm is demonstrated in Figure
\ref{fig:Outliers}. For the demonstration, we use the easting
component of displacements at a single station, SC03, which is located
on Mt. Olympus in Washington state. Station SC03 records anomalous
observations during the winter, presumably because of snow and ice
accumulation, and we want to remove these observations. The station
also records periodic westward motion from slow slip events, and we
want to keep this deformation intact. The detected outliers are shown
in Panel B. For comparison, we also show the detected outliers when we
do not include the Gaussian process $v$ in our model for the data
(Panel A). When $v$ is not included, real transient deformation
resulting from slow slip events is erroneously identified as outliers.
When $v$ is included, the identified outliers only consist of the
anomalous deformation that lacks temporal continuity.  It should be
noted that we use $\lambda = 2.5$ for this demonstration, which causes
the outlier detection algorithm to be particularly aggressive. In
Section \ref{sec:Cascadia}, we clean the data using a more tolerant
$\lambda = 4.0$.

\begin{figure*}
\includegraphics{figures/outliers_new/outliers.pdf}
\caption{
Outliers detected in the easting component of displacements at station
SC03. The orange markers indicate detected outliers. The orange line
is the best fit model to the data, which is used to compute the
residual vector $\mitbf{r}$. The model being fit to the data in Panel
A is $\G\mitbf{m}$, and the model in Panel B is $\G\mitbf{m} +
v(\points)$.
}
\label{fig:Outliers}
\end{figure*}

\bibliographystyle{gji}
\bibliography{mybib}

\end{document}