### 2017-11-09 ResponseToReviewers/ResponseToReviewers.tex
began response to reviewer

\documentclass[10pt,a4paper]{letter}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}

\begin{document}

\signature{Trever T. Hines}

\begin{letter}{}
\opening{Dear Editor,}

We were pleased with the positive comments from you and the reviewers, and we feel that the comments brought up valid concerns. We hope that we have addressed these concerns sufficiently in our revision. In order to fully address the comments, we have included a supplementary material document.  We have also made some changes to the manuscript which were not specifically requested by the reviewers.  Most of these changes were stylistic and do not change the content of the manuscript.  One of the more substantial changes was that we merged Figures 7 and 14, which are both showing observed and predicted radial components of postseismic displacements, facilitating direct comparison of the figures.

We added supporting figures S3 and S4, which break down the predicted displacements of our preferred model into an elastic and viscoelastic component.  We feel that these figures help support our discussion on lines 509 to 523, where we attribute inferences of fault slip and viscoelastic relaxation to different aspects of the observed postseismic displacements.

On line 500 to 502 of our previous version of this manuscript we make the claim that “After one year, afterslip is inferred to be deeper down on the Sierra Cucapah segment, which is describing much of the sustained near-field postseismic deformation.”  This sentence is not accurate of the preferred model we present in the manuscript. As can be clearly seen in the new figures S3-S4, afterslip is describing some of the later near-field deformation, although most of it is being described by viscoelastic relaxation.  We have reworded this passage and updated the introduction and conclusion to accurately describe our interpretations of the mechanisms driving later near-field deformation.  This does not significantly alter the main conclusions of our manuscript, because we are very clear, both in the original and revised versions, that the deformation mechanism driving later near-field deformation cannot be definitively resolved with these data.

Each of the reviewer comments are included below followed by our response and a description of how these comments were addressed in this revised manuscript.

%% REVIEWER 1
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\textbf{Reviewer \#1:}\newline

\textbf{Review of the paper : Revealing transient strain in geodetic data
with Gaussian process regression}

\textbf{Overview of the work: The paper focuses on a new non-parametric
Bayesian method to estimate transient strain rate from GNSS data. It
is based on Gaussian Process regression. The method also includes an
outlier removal processing step. This technique is then applied to the
detection of slow slip events in Cascadia. The validity of the
detected transient signals is accomplished via comparison with the
seismic tremor recorded along the Cascadia range.}

\textbf{The article is well-written and your algorithm is very interesting. I
think the use of Gaussian Process Regression is novel and open new
roads in GNSS data analysis. I believe that you spend a lot of efforts
in trying to justify most of the assumptions behind your model,
together with the effects of using different kind of prior. Perhaps,
that overshadows slightly your results. You could consider to
reorganize the article to make it more readable for geophysicists.}

\textbf{I also require that a comparison between your outlier removal model
and one other algorithm could help to support your claims about the
efficiency of your method.}

\textbf{Otherwise, I have underlined a few clarifications needed about the
stochastic model and some paragraphs within the text.}

\textbf{General Comments:}

\textbf{About the method:}

\textbf{The method is well established. The assumptions to justify the various
hypothesis on the deterministic models of the GNSS time series are
generally well supported (seasonal signal, tectonic rate, offsets).}

\textbf{1- Now, the correlation of the stochastic processes is not completely
clear. When the authors justify the model behind equation 3, wij is a
normally distributed, uncorrelated noise. In statistic, it is equal to
a WGN process. Then, the authors introduce a separate parameter (eta)
which models the temporally correlated noise. This parameter is also
following a Gaussian distribution, with a zero-mean and a covariance
matrix $C_{\eta}$ (i.e. line 10 p5). This approach is not common in
modelling GNSS time series. The standard is based on Williams 2003,
where the author established a stochastic model with a covariance
matrix as sum of identity matrix for the white noise and another
matrix which represents the coloured noise. The coloured noise
covariance matrix is defined differently (see the literature – Bock and
Melgar Reviews of Geophysics 2016, He et al., Journal of geodynamics
2017).}

\textbf{2- First, the author should give a summary of stochastic noise
modelling in GNSS. I also think that a discussion is needed to relate
to previous models. Why using separated stochastic models? Also I
would comment on the possibility that the (low) spatial correlation
between the parameters in your deterministic model (i.e. secular
velocities) could be absorbed in the estimation of the covariance
matrix $C_{\eta}$.}

\textbf{Now, I am confused When looking at p.8 Section 4.1, the description of
the noise models should be given as a subsection of Section 2 p.4. I
would also improve the literature review. Several other models have
also been discussed (i.e. Montillet et al., 2014 uses a fractional
Brownian motion model; ARMA, ARFIMA, GGM or Band pass noise – see Het
et al., 2017).}

\textbf{Outlier detection:}

\textbf{1- P7 : I would give a longer summary in the introduction about all
the efforts in outlier detection in GNSS time series. The first
paragraph p.7 is not enough.}

\textbf{2- Also, the paper would be improved if the proposed outlier detection
method is compared with existing ones such as the Hector software
package (Bos et al., 2013). For example, the comparison between this
method and another one could be done when applying the algorithm in p.
13 (last paragraph below equation 18).}

\textbf{Results:}

\textbf{The results are generally well explained. However, the use of
different priors and the comparison in the text is sometimes
confusing. Perhaps, a table could summary the different priors and the
main results. It would ease the reading of Section 4.2, 4.3 and then
the discussion in Section 5.}

\textbf{Minor issues:}

\textbf{P2 line 30 : “fidelity” replace with “reliability”.}

\textbf{P2 line 30-31 “ Developing and improving upon methods for deriving
secular...area of research.” Need references.}

\textbf{p.2 line 43 :”too large of an area” ... perhaps “ a very large area”}

\textbf{P5 line 32 : “formal data uncertainty”: can you define it?}

\textbf{P 6 : Need reference to show Equation 15.}

\textbf{P 7 line 54-60: This paragraph can be further enhanced with references
(Bock and Melgar, 2016, Gazeaux et al., 2013).}

\textbf{P 8 : first paragraph. add references on Cascadia range (Aguiar,
Melbourne, Scriver 2009), (Melbourne and Szeliga 2005), (Melbourne and
Webb, 2003, 2002).}

\textbf{P.8 line 20 delete www.unavco.org (only in Acknowledgment)}

\textbf{p. 11 line 45 “common mode error” you need to add some references and
define it properly.}

\textbf{p.13 : “Wendland functions have compact support and hence their
corresponding covariance.. sparse”. I would just mention that the
associated covariance matrices are sparse.}

\textbf{p.13 line 22 what is an “isotropic Gaussian process”? please add a
definition.}

\textbf{P.20 : “ Using a compact covariance function ..” Again I would just
say that the covariance matrix is sparse and that s why you have the
mathematical simplification.}

\textbf{p.20 line 33 “ computational burden” I think that refers to p.20 line
60 “prohibitive when using sveral years of daily GNSS”. It would
interesting to have some number. How do you quantify the computation
time to process longer and longer time series?}

%% REVIEWER 2
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\textbf{Reviewer \#2:}\newline

\textbf{Comments to the Author(s)}
\textbf{Hines/Hetland GJI Review 08/2017}

\textbf{This is a well written paper and I have just a few minor
clarifications and suggestions below.  One point that might be of
interest to address is would it be possible to “approximate” this
approach so that it can run much faster?  I gather from the text that
the method is intensive for CPU and memory use.}

\textbf{Page 3: Last line: min(t,t’) needs as scaling factor to make units for
covariance e.g., m$^2$/yr}

\textbf{Page 4: Stochastic model: It would seem that the assumption of no
covariance between the north and east components is not a very good
one since there is a tectonic framework in which the transients occur
and its unlikely this framework aligns along the Cardinal directions.
Maybe add a statement of impact of the neglect and possibly the idea
of reorienting the “axes” to align with the tectonic framework e.g.,
perpendicular to the subduction zone interface in this case?}

\textbf{Page 4/5 Equation (3): Is it worth discussing at this point whether
the estimates of u and n (eta) can be separated without explicit prior
knowledge (which we may not have).  For example, orbit modeling errors
on one satellite (e.g., due to an unknown yaw problem in the
spacecraft) wilt generate a spatially and temporally correlated error
in $d_{ij}$.  How does this (stochastically non-modeled) error not project
in the u estimates?}

\textbf{Page 6: Equation 10: Maybe it is discussed later in the paper but it
is probably worthwhile stating at this point how the inverse is
performed with a block 0 in the matrix?}

\textbf{Page 6: Equation 12: Maybe some additional explanation is needed here.
How are these partial derivatives formed.  My understanding the u
estimates at this point correspond one-to-one with the positions and
times of the GNSS position determinations.  Did I miss something?}

\textbf{Page 8: Partly addresses issue raised about eqn 3.  It probably should
be noted that wether conditions east of -121 deg longitude are
different to the coast so there could be possible problems with this
noise model.}

%% Editor Review
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\textbf{Editor:}\newline

\textbf{Editor’s Review of GJI MS 17.0593 ‘‘Revealing transient strain...’’ by
Hines and Hetland This is an novel algorithm applied to a problem of
great current interest. Based on the reviewer’s comments and my own
reading, I am recommending moderate revision. The authors’ response
should include a description of their changes, and also a version of
the paper on which these can be seen (just as highlights in the text).}

\textbf{Many of the reviewers’ comments address places where the paper could
be more clearly written, and my overall comment would be the same. The
current version perhaps spends a little too much time on the
background and not enough on explaining the method and its derivation.
As shown by the references below (these include some the authors
cite), and the Dermanis paper, which I have also attached, strain
estimation from scattered data is an area in which there has been a
lot more prior work than the authors mention. I am including these
references not because the authors need to cite many or even most of
them, but only to make the point that the problem of estimating
strain, and of looking for transients, is a familiar one: for many
readers, myself among them, much more familiar than the statistical
methods applied here. So a slightly fuller discussion of these methods
(perhaps a little more background and some reminders of, eg, what a
hyperparameter is) would be welcome to many readers; and increased
clarity of presentation usu- ally means that the paper will be used
and cited more.}

\textbf{I feel the same way about the section on detecting outliers: again, a
well-worked area. Iter- ative fitting and outlier detection, as an
overall strategy, hav e been used for a long time. I appreci- ate that
by imposing spatial and temporal conditions you can identify outliers
not otherwise obvi- ous, but it would help to show an example of this,
rather than the current Figure 4, in which most of the identified
cases are ones that any method could easily find.}

\textbf{A few specifics (numbers are page+line):}

\textbf{2+26: possibility not risk: the latter is a technical term in seismic
hazard.}

\textbf{2+47-48: I do not think this description of Shen’s method (which I’ve
used and programmed) is right, since it allows the deformation
gradients to vary in space. I’ve always thought of it as a kind of
adaptive kernel smoother, somewhat like applying 2-D loess; so it is
only depen- dent on nearby points, and has no assumption of uniformity
anywhere. (Actually this raises a more general question: is the method
given here use local or global support?)}

\textbf{3+7: could also}

\textbf{3+32: redo the sentence; as it stands, you are saying that the SCEC
exercise calculates strain rates.}

\textbf{3+37: a geophysical signal}

\textbf{3+47: function not signal}

\textbf{3+56: by a Gaussian, not with}

\textbf{4+37: displacement (not plural)}

\textbf{6+56: I suspect that equation (15) took some effort to derive; could
some details go into an Appendix?}

\textbf{Figure 6: I like that the strain uncertainties can be shown, but is
there a reason not to plot principle strains using arrows, showing
errors by an ellipse around the tips?}

\closing{Thank you for your continued consideration,}

\end{letter}
\end{document}