### 2017-11-10 ResponseToReviewers/ResponseToReviewers.tex
did more

\documentclass[10pt,a4paper]{letter}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}

\begin{document}

\signature{Trever T. Hines}

\begin{letter}{}
\opening{Dear Editor,}

We were pleased with the positive comments from you and the reviewers,
and we feel that the comments brought up valid concerns. We hope that
we have addressed these concerns sufficiently in our revision. In
order to fully address the comments, we have included a supplementary
material document.  We have also made some changes to the manuscript
which were not specifically requested by the reviewers.  Most of these
changes were stylistic and do not change the content of the
manuscript.  One of the more substantial changes was that we merged
Figures 7 and 14, which are both showing observed and predicted radial
components of postseismic displacements, facilitating direct
comparison of the figures.

We added supporting figures S3 and S4, which break down the predicted
displacements of our preferred model into an elastic and viscoelastic
component.  We feel that these figures help support our discussion on
lines 509 to 523, where we attribute inferences of fault slip and
viscoelastic relaxation to different aspects of the observed
postseismic displacements.

On line 500 to 502 of our previous version of this manuscript we make
the claim that “After one year, afterslip is inferred to be deeper
down on the Sierra Cucapah segment, which is describing much of the
sustained near-field postseismic deformation.” This sentence is not
accurate of the preferred model we present in the manuscript. As can
be clearly seen in the new figures S3-S4, afterslip is describing some
of the later near-field deformation, although most of it is being
described by viscoelastic relaxation.  We have reworded this passage
and updated the introduction and conclusion to accurately describe our
interpretations of the mechanisms driving later near-field
deformation.  This does not significantly alter the main conclusions
of our manuscript, because we are very clear, both in the original and
revised versions, that the deformation mechanism driving later
near-field deformation cannot be definitively resolved with these
data.

Each of the reviewer comments are included below followed by our
response and a description of how these comments were addressed in
this revised manuscript.

%% REVIEWER 1
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\textbf{Reviewer \#1:}\newline

\textbf{Review of the paper : Revealing transient strain in geodetic data
with Gaussian process regression}

\textbf{Overview of the work: The paper focuses on a new non-parametric
Bayesian method to estimate transient strain rate from GNSS data. It
is based on Gaussian Process regression. The method also includes an
outlier removal processing step. This technique is then applied to the
detection of slow slip events in Cascadia. The validity of the
detected transient signals is accomplished via comparison with the
seismic tremor recorded along the Cascadia range.}

\textbf{The article is well-written and your algorithm is very interesting. I
think the use of Gaussian Process Regression is novel and open new
roads in GNSS data analysis. I believe that you spend a lot of efforts
in trying to justify most of the assumptions behind your model,
together with the effects of using different kind of prior. Perhaps,
that overshadows slightly your results. You could consider to
reorganize the article to make it more readable for geophysicists.}

I have taken several steps to make my revised draft more accessible to
someone who does not have a strong background in stochastic processes:
I rewritten the methods section to be clearer, contain more examples,
and elaborate on potentially unfamiliar concepts. I have also added
more references to the methods section to make it easier for the
reader to find additional information. I sincerely hope that my
revised draft is more readable to geophysicists.

\textbf{I also require that a comparison between your outlier removal model
and one other algorithm could help to support your claims about the
efficiency of your method.}

(See my response to the comment ``Also, the paper would be improved if
the proposed outlier detection ...")

\textbf{Otherwise, I have underlined a few clarifications needed about the
stochastic model and some paragraphs within the text.}

\textbf{General Comments:}

\textbf{About the method:}

\textbf{The method is well established. The assumptions to justify the various
hypothesis on the deterministic models of the GNSS time series are
generally well supported (seasonal signal, tectonic rate, offsets).}

\textbf{1- Now, the correlation of the stochastic processes is not completely
clear. When the authors justify the model behind equation 3, wij is a
normally distributed, uncorrelated noise. In statistic, it is equal to
a WGN process. Then, the authors introduce a separate parameter (eta)
which models the temporally correlated noise. This parameter is also
following a Gaussian distribution, with a zero-mean and a covariance
matrix $C_{\eta}$ (i.e. line 10 p5). This approach is not common in
modelling GNSS time series. The standard is based on Williams 2003,
where the author established a stochastic model with a covariance
matrix as sum of identity matrix for the white noise and another
matrix which represents the coloured noise. The coloured noise
covariance matrix is defined differently (see the literature – Bock and
Melgar Reviews of Geophysics 2016, He et al., Journal of geodynamics
2017).}

I have changed my description of the data (eq. 3 in the revised draft)
so that it includes a single term, $\eta$, that encompasses correlated
and uncorrelated noise. I then break $\eta$ down into correlated and
uncorrelated components in Section 3.1 (eq. 25). The covariance matrix
for $\eta$ in eq. 26 is more consistent with the literature.

\textbf{2- First, the author should give a summary of stochastic noise
modelling in GNSS. I also think that a discussion is needed to relate
to previous models. Why using separated stochastic models? Also I
would comment on the possibility that the (low) spatial correlation
between the parameters in your deterministic model (i.e. secular
velocities) could be absorbed in the estimation of the covariance
matrix $C_{\eta}$.}

I believe the reviewer is refering to page 4-5 of the original draft,
where I first introduce $\eta$. I do not think this is an appropriate
place for a discussion on the noise model, or for a summary of past
studies. The paper is laid out so that Section 2 contains a general
description of the method for estimating transient strain rates, and
Section 3 contains details that are specific to the particular
application. There is no universally appropriate noise model so I
decided to save a discussion on the noise model for Section 3. This is
clarified in Section 2 of the revised draft when I say "The
appropriate noise model may vary depending on the application, and we
hold off on specifying the covariance matrix, $\mitbf{C}_{\eta_i}$,
until Section \ref{sec:NoiseModel}."

\textbf{Now, I am confused When looking at p.8 Section 4.1, the
description of the noise models should be given as a subsection of
Section 2 p.4. I would also improve the literature review. Several
other models have also been discussed (i.e. Montillet et al., 2014
uses a fractional Brownian motion model; ARMA, ARFIMA, GGM or Band
pass noise – see Het et al., 2017).}

Again, we think that the description of the noise model should remain
in Section 3 because it is specific to the Pacific Northwest
application.

We have added a few more references in Section 3.1 of the revised
draft TODO to emphasize that GNSS noise modeling is a well-worked
area. We refer the reader to Bock and Melgar 2016 and He et al 2017
for a more thorough review.

\textbf{Outlier detection:}

\textbf{1- P7 : I would give a longer summary in the introduction about all
the efforts in outlier detection in GNSS time series. The first
paragraph p.7 is not enough.}

In Section 2.2 of the revised draft, I have expanded the discussion of
previous efforts to detect outliers in GNSS data.

\textbf{2- Also, the paper would be improved if the proposed outlier detection
method is compared with existing ones such as the Hector software
package (Bos et al., 2013). For example, the comparison between this
method and another one could be done when applying the algorithm in p.
13 (last paragraph below equation 18).}

My outlier detection algorithm is quite similar to what is used in Bos
et al 2013 (and many other papers), in that the outliers are
determined based on the residuals for a best fitting model. The only
difference is that I am fitting the data with a stochastic model
rather than a parametric model consisting of a linear trend and
sinusoids. When using a parametric model, I found that deformation
from slow slip events would sometimes be erroneously identified as an
outlier. The stochastic model used in my outlier detection method is
more flexible and able to describe this transient deformation, and so
only high frequency anomalous observations get identified as outliers.
I have included Figure A1 to demonstrate this.

Of course, a parametric model that includes, for example, B-splines
would be just as effective at detecting outliers without removing
geophysical signal. So I am not suggesting that my outlier detection
method is a significant advancement or even a novel method. Instead, I
am only describing my outlier detection method for the sake of
completeness. The previous version of my paper seems to have put too
much emphasis on the outlier detection method, when the main focus
should remain on the method for estimating transient strain rates. To
help keep my paper focused, I placed the description of the outlier
detection method in the appendix.

\textbf{Results:}

\textbf{The results are generally well explained. However, the use of
different priors and the comparison in the text is sometimes
confusing. Perhaps, a table could summary the different priors and the
main results. It would ease the reading of Section 4.2, 4.3 and then
the discussion in Section 5.}

In the previous draft, I was discussing two different priors: a prior
for the outlier detection algorithm, and a prior used in computing
transient strain rates. I acknowledge that this was convoluted. In the
revised draft, all the details of the outlier detection algorithm are
confined to the appendix. I hope this clarifies any confusion.

\textbf{Minor issues:}

\textbf{P2 line 30 : “fidelity” replace with “reliability”.}

Done

\textbf{P2 line 30-31 “ Developing and improving upon methods for deriving
secular...area of research.” Need references.}

TODO

\textbf{p.2 line 43 :”too large of an area” ... perhaps “ a very large area”}

Done

\textbf{P5 line 32 : “formal data uncertainty”: can you define it?}

By "formal data uncertainties" I meant "the derived uncertainties that
accompany the GNSS displacement solutions". The revised text has been
clarified.

\textbf{P 6 : Need reference to show Equation 15.}

I have added a few more steps in the derivation of eq. 15.

\textbf{P 7 line 54-60: This paragraph can be further enhanced with references
(Bock and Melgar, 2016, Gazeaux et al., 2013).}

I have added a reference to Gazeaux et al. 2013.

\textbf{P 8 : first paragraph. add references on Cascadia range (Aguiar,
Melbourne, Scriver 2009), (Melbourne and Szeliga 2005), (Melbourne and
Webb, 2003, 2002).}

We agree that we need to acknowledge the extensive amount of research
on Cascadia SSEs. In the revised draft, we refer the reader to the
review by Schartz and Rokosky 2007.

\textbf{P.8 line 20 delete www.unavco.org (only in Acknowledgment)}

Done

\textbf{p. 11 line 45 “common mode error” you need to add some references and
define it properly.}

In Section 3.1 of the revised draft, where we elaborate on the noise
model, we describe common mode error by saying "Another significant
source of noise in GNSS data is common mode error (e.g., Wdowinski et
al. 1997; Dong et al. 2006), which is noise that is highly spatially
correlated." We believe this definition and these references are
sufficient.

\textbf{p.13 : “Wendland functions have compact support and hence their
corresponding covariance.. sparse”. I would just mention that the
associated covariance matrices are sparse.}

I recognize that it may be a bit redundant to say that the covariance
function is compact and thus the covariance matrices are sparse.
However, Wendland covariance functions are typically described as
"compact" in the literature, and I think it is a key adjective that
should not be dropped out. I have left the wording as is.

\textbf{p.13 line 22 what is an “isotropic Gaussian process”? please add a
definition.}

An Gaussian process is isotropic if its covariance function,
$C(x,x')$, can be written as a function of $||x - x'||_2^2$. An
isotropic Gaussian process is also stationary, meaning that its
statistical properties are invariant to translations. We have replace
"isotropic" with "stationary" in the text because "stationary" is a
more appropriate, and perhaps more familiar, term.

\textbf{P.20 : “ Using a compact covariance function ..” Again I would just
say that the covariance matrix is sparse and that s why you have the
mathematical simplification.}

My personal preference is to leave the wording as is.

\textbf{p.20 line 33 “ computational burden” I think that refers to p.20 line
60 “prohibitive when using sveral years of daily GNSS”. It would
interesting to have some number. How do you quantify the computation
time to process longer and longer time series?}

I agree that numerical values would be interesting; however, I am
reluctant to quantify the computational cost of evaluation transient
strain rates. There are many factors that influence the computational
cost (whether sparse matrices are used, the type of sparse matrices
that are used, the linear solver algorithm, etc.) and I am not
convinced that my implementation is the most efficient. So I do not
give details on the computation time, but I note in the discussion
that this is an area for further research.

%% REVIEWER 2
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\textbf{Reviewer \#2:}\newline

\textbf{Comments to the Author(s)}
\textbf{Hines/Hetland GJI Review 08/2017}

\textbf{This is a well written paper and I have just a few minor
clarifications and suggestions below.  One point that might be of
interest to address is would it be possible to “approximate” this
approach so that it can run much faster?  I gather from the text that
the method is intensive for CPU and memory use.}

\textbf{Page 3: Last line: min(t,t’) needs as scaling factor to make units for
covariance e.g., m$^2$/yr}

\textbf{Page 4: Stochastic model: It would seem that the assumption of no
covariance between the north and east components is not a very good
one since there is a tectonic framework in which the transients occur
and its unlikely this framework aligns along the Cardinal directions.
Maybe add a statement of impact of the neglect and possibly the idea
of reorienting the “axes” to align with the tectonic framework e.g.,
perpendicular to the subduction zone interface in this case?}

\textbf{Page 4/5 Equation (3): Is it worth discussing at this point whether
the estimates of u and n (eta) can be separated without explicit prior
knowledge (which we may not have).  For example, orbit modeling errors
on one satellite (e.g., due to an unknown yaw problem in the
spacecraft) wilt generate a spatially and temporally correlated error
in $d_{ij}$.  How does this (stochastically non-modeled) error not project
in the u estimates?}

\textbf{Page 6: Equation 10: Maybe it is discussed later in the paper but it
is probably worthwhile stating at this point how the inverse is
performed with a block 0 in the matrix?}

\textbf{Page 6: Equation 12: Maybe some additional explanation is needed here.
How are these partial derivatives formed.  My understanding the u
estimates at this point correspond one-to-one with the positions and
times of the GNSS position determinations.  Did I miss something?}

\textbf{Page 8: Partly addresses issue raised about eqn 3.  It probably should
be noted that wether conditions east of -121 deg longitude are
different to the coast so there could be possible problems with this
noise model.}

%% Editor Review
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\textbf{Editor:}\newline

\textbf{Editor’s Review of GJI MS 17.0593 ‘‘Revealing transient strain...’’ by
Hines and Hetland This is an novel algorithm applied to a problem of
great current interest. Based on the reviewer’s comments and my own
reading, I am recommending moderate revision. The authors’ response
should include a description of their changes, and also a version of
the paper on which these can be seen (just as highlights in the text).}

\textbf{Many of the reviewers’ comments address places where the paper could
be more clearly written, and my overall comment would be the same. The
current version perhaps spends a little too much time on the
background and not enough on explaining the method and its derivation.
As shown by the references below (these include some the authors
cite), and the Dermanis paper, which I have also attached, strain
estimation from scattered data is an area in which there has been a
lot more prior work than the authors mention. I am including these
references not because the authors need to cite many or even most of
them, but only to make the point that the problem of estimating
strain, and of looking for transients, is a familiar one: for many
readers, myself among them, much more familiar than the statistical
methods applied here. So a slightly fuller discussion of these methods
(perhaps a little more background and some reminders of, eg, what a
hyperparameter is) would be welcome to many readers; and increased
clarity of presentation usu- ally means that the paper will be used
and cited more.}

\textbf{I feel the same way about the section on detecting outliers: again, a
well-worked area. Iter- ative fitting and outlier detection, as an
overall strategy, hav e been used for a long time. I appreci- ate that
by imposing spatial and temporal conditions you can identify outliers
not otherwise obvi- ous, but it would help to show an example of this,
rather than the current Figure 4, in which most of the identified
cases are ones that any method could easily find.}

\textbf{A few specifics (numbers are page+line):}

\textbf{2+26: possibility not risk: the latter is a technical term in seismic
hazard.}

\textbf{2+47-48: I do not think this description of Shen’s method (which I’ve
used and programmed) is right, since it allows the deformation
gradients to vary in space. I’ve always thought of it as a kind of
adaptive kernel smoother, somewhat like applying 2-D loess; so it is
only depen- dent on nearby points, and has no assumption of uniformity
anywhere. (Actually this raises a more general question: is the method
given here use local or global support?)}

\textbf{3+7: could also}

\textbf{3+32: redo the sentence; as it stands, you are saying that the SCEC
exercise calculates strain rates.}

\textbf{3+37: a geophysical signal}

\textbf{3+47: function not signal}

\textbf{3+56: by a Gaussian, not with}

\textbf{4+37: displacement (not plural)}

\textbf{6+56: I suspect that equation (15) took some effort to derive; could
some details go into an Appendix?}

\textbf{Figure 6: I like that the strain uncertainties can be shown, but is
there a reason not to plot principle strains using arrows, showing
errors by an ellipse around the tips?}

\closing{Thank you for your continued consideration,}

\end{letter}
\end{document}